Compare consensus cell types with SCimilarity to consensus cell types without SCimilarity

### 2025-09-12 analyses/cell-type-consensus/exploratory-notebooks/07-compare-consensus.Rmd
old is new and new is old

# first add columns for total cells per library, number of new and old cell types
grouped_df <- all_results_df |> 
    dplyr::group_by(library_id) |> 
    dplyr::mutate(
      total_cells_per_library = dplyr::n(),
      num_old_celltypes = length(unique(consensus_annotation)),
      num_new_celltypes = length(unique(existing_consensus_celltype_annotation))
    ) |>
    dplyr::ungroup()

# get stats for old consensus cell types
old_consensus_df <- grouped_df |> 
    dplyr::group_by(project_id, project_label, diagnosis, library_id, sample_type, existing_consensus_celltype_annotation, existing_consensus_celltype_ontology) |> 
    dplyr::summarize(num_celltypes = unique(num_new_celltypes), # constant for each library
                     total_cells_per_annotation = dplyr::n(),
                     total_cells_per_library = unique(total_cells_per_library)) |>
    dplyr::mutate(
      # add percentage 
      percent_cells_annotation = round((total_cells_per_annotation / total_cells_per_library) * 100, 2)
    ) |> 
    dplyr::ungroup()

# stats for new consensus cell types
new_consensus_df <- grouped_df |> 
    dplyr::group_by(project_id, project_label, diagnosis, library_id, sample_type, consensus_annotation, consensus_ontology) |> 
    dplyr::summarize(num_celltypes = unique(num_old_celltypes), # constant for each library
                     total_cells_per_annotation = dplyr::n(),
                     total_cells_per_library = unique(total_cells_per_library)) |>
    dplyr::mutate(
      # add percentage 
      percent_cells_annotation = round((total_cells_per_annotation / total_cells_per_library) * 100, 2)
    ) |> 
    dplyr::ungroup()

>>>>

# first add columns for total cells per library, number of new and old cell types
grouped_df <- all_results_df |> 
    dplyr::group_by(library_id) |> 
    dplyr::mutate(
      total_cells_per_library = dplyr::n(),
      num_old_celltypes = length(unique(existing_consensus_celltype_annotation)),
      num_new_celltypes = length(unique(consensus_annotation))
    ) |>
    dplyr::ungroup()

# get stats for old consensus cell types
old_consensus_df <- grouped_df |> 
    dplyr::group_by(project_id, project_label, diagnosis, library_id, sample_type, existing_consensus_celltype_annotation, existing_consensus_celltype_ontology) |> 
    dplyr::summarize(num_celltypes = unique(num_old_celltypes), # constant for each library
                     total_cells_per_annotation = dplyr::n(),
                     total_cells_per_library = unique(total_cells_per_library)) |>
    dplyr::mutate(
      # add percentage 
      percent_cells_annotation = round((total_cells_per_annotation / total_cells_per_library) * 100, 2)
    ) |> 
    dplyr::ungroup()

# stats for new consensus cell types
new_consensus_df <- grouped_df |> 
    dplyr::group_by(project_id, project_label, diagnosis, library_id, sample_type, consensus_annotation, consensus_ontology) |> 
    dplyr::summarize(num_celltypes = unique(num_new_celltypes), # constant for each library
                     total_cells_per_annotation = dplyr::n(),
                     total_cells_per_library = unique(total_cells_per_library)) |>
    dplyr::mutate(
      # add percentage 
      percent_cells_annotation = round((total_cells_per_annotation / total_cells_per_library) * 100, 2)
    ) |> 
    dplyr::ungroup()
